\section{The Method of Containers}

\subsection{Some nice problems}
There are many reasons to understand the method of hypergraph
containers. Here are some cool problems we may be interested
in.
\begin{problem}
	\label{containers:prob:extremalK3}
	(Extremal questions in $G(n,p)$). We define
	\[
		\text{ex}(G,H) := \max \{ e(G'): G' \subset G,\; G \text{ is }
		H-\text{free}\}.
	\]
	Note that $\text{ex}(n,H) = \text{ex}(K_n,H)$. What can we say about
	$\text{ex}(G(n,p),K_3)$?
\end{problem}

Some easy bounds can be found. Given any sampling of $G(n,p)$, we can
always find a bipartite graph with half the edges. Similarly
another strategy is just to kill off triangles by removing an edge
from each. This already gives
\[
	\text{ex}(G(n,p), K_3) \geq
	\begin{cases}
		e(G(n,p))/2\\
		e(G(n,p)) - k_3(G(n,p))
	\end{cases}
\]
by finding the best regime for $p$ in each we get
\[
	\text{ex}(G(n,p), K_3) \geq
	\begin{cases}
		pn^2/4 \quad \text{for any}\,p \in [0,1]\\
		(1 - o(1))pn^2/2 \quad \text{for}\, p \ll \frac{1}{\sqrt{n}}
	\end{cases}
\]
it turns out this is sharp.

\begin{problem}
	When does $G(n,p) \to K_3$? Define
	\[
		p^*(n) = \inf_{p \in [0,1]} \{p : \Pr(G(n,p) \to K_3) \geq 1/2\}.
	\]
	can we find this threshold?
\end{problem}
We know that if $G(n,p)$ contains a $K_6$ we are done by Ramsey so
$p^* \lesssim n^{-2/5}$. But is this it?
Should avoiding $K_6$ be enough or is there some deeper more complex
structure - the answer is obviously yes.

\begin{problem}
	Let $A \subset [n]_p$ with $|A| \geq \delta np$, does $A$ contain a $3$-AP?
\end{problem}
Doing deletion we get that it is true if  $AP_3([n]_p) = o(np)$. We
know $AP_3([n]_p) \eqsim n^2p^3$, so if $p \ll 1/\sqrt{n}$
we get the result. What happens if $p \gg 1/\sqrt{n}$?

Very related to \ref{containers:prob:countingk3} is question:
\begin{problem}
	\label{containers:prob:countingk3}
	How many $K_3$-free graphs on $n$ vertices with $m$ edges are there?
\end{problem}
If we call this quantity $X(n,m)$, we know (because we can always
pick from a bipartite graph) that
\[
	X(n,m) \geq \binom{n^2/4}{m}.
\]
If $m$ is sufficiently low, we can use $G(n,p)$ for our advantage! We know that
\[
	\Pr(K_3 \not \subset G(n,p)) \eqsim \exp(-n^3p^3)
\]
if $p \ll n^{-1/2}$ by Janson and Harris inequality. Noting that $G(n,
m) \eqsim G(n,p)$ with $p = m/\binom{n}{2}$,
we get something that says
\[
	X(n,m) \eqsim \exp(-m^3/n^3)\binom{\binom{n}{2}}{m} \quad \text{if
	}\, m \ll n^{3/2}.
\]
Is the first bound optimal?

\subsection{Attempts at handling them}

Let's follow a naive approach for tackling
\ref{containers:prob:extremalK3}. So let's see first moment.
Let
\[Y = \#\big\{G: K_3 \not\subset G \subset G(n,p)\; \text{ with }
e(G) > m = (1/2 + \eps)p\binom{n}{2} \big\}\]
the number of such bad subgraphs of $G(n,p)$. We would like to show
$Y = 0$ with high probability if $p \gg 1/\sqrt{n}$. If we calculate
$\Ex Y$ we get
\[
	\Ex Y = \sum_{\substack{K_3 \not \subset G\\e(G) = m}} \Pr(G \subset
	G(n,p)) \geq \binom{n^2/4}{m} p^m \eqsim
	\bigg(\frac{en^2p}{4m}\bigg)^m \eqsim e^m
\]
where for the lowerbound we only considered the subsets of a single
bipartite graph. But $\Ex Y$ is \textbf{HUGE}, so can we better this?

A very interesting point made by Rob is to look where we are losing
here. Every once and while, when you look at an equation and it
explodes, it is a good idea to see why.
Maybe we could handle the bad case differently.

Here we are losing a
ton because there are many subsets of the bipartite graph, but this
seems a bit unrealistic,
a bipartite graph contains at most $1/2$ of the edges of $G(n,p)$ and
here we are asking for an $\eps$ deviation.
To be more precise, let's check another thing.

What can we say about bipartite $H$ in $G(n,p)$ with $m$ edges? Can
we find many of those?
Note that every bipartite graph $H$ is a subset of a complete
bipartite $G$ which has at most $n^2/4$ edges. So we could say
\[
	\Pr(\exists\; \text{bip.}\,H \subset G(n,p)\; \text{with } e(H) = m) =
	\sum_{\substack{K_{(n/2,n/2)} \sim G\\G \subset G(n,p)}}
	\Pr(\exists\; \text{bip.}\,H \subset G \cap G(n,p)\; \text{with } e(H) = m)
\]
It might look that we made our lives worse, but in reality this is a
way simpler thing. For such $H$ to exists, $e(G \cap G(n,p)) \geq m$, so
here we can already apply a deviation bound. Furthermore, the number
of complete bipartites is $2^{n-1}$ so applying the union bound seems
like a great idea.
We get (using \ref{lemma:expec_binomial_bounds})
\begin{align*}
	\sum_{\substack{K_{(n/2,n/2)} \sim G\\G \subset G(n,p)}}
	\Pr(\exists\; \text{bip.}\,H \subset G \cap G(n,p)\; \text{with }
	e(H) = m) &\leq
	2^n\Pr(\text{Bin}(n^2/4,p) \geq m)\\
	&\leq 2^n\exp(-\eps^2n^2p/4) \to 0
\end{align*}
if $p \gg 1/n$ so we win by miles in this bipartite case.

What we actually did here is a container theorem for bipartite graphs.
Let $\mathcal{G}$ be the family of complete bipartite graphs on $n$ vtxs,
we know that
\begin{enumerate}[label=(\alph*)]
	\item $|\mathcal{G}|$ is small $\leq 2^n$.
	\item Every $G \in \mathcal{G}$ is bipartite.
	\item For every bipartite $H \subset K_n$, there exists $G \in
		\mathcal{G}$ such that $H \subset G$. ($G$ is the container for $H$).
\end{enumerate}
This family allowed us to do way more efficient union bound. And this
is what will try to do to tackle
\ref{containers:prob:extremalK3}.

For our problem, this we need another container theorem which we will
prove later:
\begin{theorem}
	\label{trm:container-k3-free-bad}
	(Container theorem for $K_3$-free graphs). There exists a family
	$\mathcal{G}$ of graphs on $n$-vtxs, s.t
	\begin{enumerate}[label=(\alph*)]
		\item $\mathcal{G}$ is small, $|\mathcal{G}| \leq 2^{n^{3/2}(\log n)^c}$.
		\item Every $G \in \mathcal{G}$ has "few" triangles, $o(n^3)$.
		\item For every $K_3$-free graph $H$, there exists a container $G
			\in \mathcal{G}$ s.t. $H \subset G$.
	\end{enumerate}
\end{theorem}

\begin{proof}[An answer for \ref{containers:prob:extremalK3}]
	Note that we can upperbound $\Pr(\exists K_3 \not \subset H \subset
	G(n,p), e(H) = m)$ using our containers.
	How many edges can our containers have? Because they have $o(n^3)$
	triangles, by supersaturation they must have at most $(1/2 +
	o(1))\binom{n}{2}$ edges.
	So following the same calculation as before (doing the union bound
	and applying our deviation result), we get
	\[
		\Pr(\exists K_3 \not \subset H \subset G(n,p), e(H) = m) \leq
		\sum_{G \in \mathcal{G}} \Pr(e(G \cap G(n,p)) \geq m) \leq
		2^{n^{3/2}\log(n)^c}e^{-\eps pn^2} \to 0
	\]
	if $p \gg \log(n)^c/\sqrt{n}$.
\end{proof}

Next time we will try to prove Theorem \ref{trm:container-k3-free-bad},
and we will do so in greater generality. We will do it for
$3$-uniform hypergraphs and apply
it for a "triangle encoding" hypergraph.
\begin{definition}
	The triangle enconding hypergraph $\mathcal{H}_{K_3}$ is a 3-uniform
	hypergraph with $V(\mathcal{H}_{k_3}) = E(K_n)$ and
	$E(\mathcal{H}_{k_3}) = \{\{e,f,g\} : e,f,g \in E(K_n) \text{ form a
	triangle}\}$.
\end{definition}

\subsection{A general hypergraph container statement with fingerprints}
%STATE GENERAL HYPERGRAPH CONTAINERS
%PROVE THE CONTAINER THEOREM FOR TRIANGLE ENCODING GRAPHS

\begin{lemma}
	\label{lemma:hypergraph_containers}
	(Hypergraph Container Lemma) For all $c \geq 1$ and $k \in \N$,
	there exists $\delta \sim 1/ck^4$ s.t the following holds.
	Let $\mathcal{H}$ be a $k$-uniform hypergraph on $n$-vtxs and $\tau
	> 0$ satisfying
	\[
		\Delta_l(\mathcal{H}) \leq
		c\tau^{l-1}\frac{e(\mathcal{H})}{v(\mathcal{H})} = c\tau^{l-1}d
	\]
	then there exists families $\mathcal{S} \subset
	\mathcal{I}(\mathcal{H})$ and $\mathcal{C}\subset P(V(\mathcal{H}))$
	and a function $f: \mathcal{S} \to \mathcal{C}$ such that
	\begin{enumerate}[label=(\alph*)]
		\item For every $S \in \mathcal{S}$, $|S| \leq \tau n$.
		\item For every $C \in \mathcal{C}$, $|C| \leq (1-\delta)n$.
		\item For every $I \in \mathcal{I}(\mathcal{H})$, there exists
			$S(I) \in \mathcal{S}$ s.t
			\[
				S(I) \subset I \subset C(I) = f(S(I)).
			\]
	\end{enumerate}
\end{lemma}

With its full generality, we can prove the container theorem for
$K_3$ encoding graphs.
\begin{theorem}
	\label{trm:container_for_triangles}
	(Containers for triangle free graphs) For every $\eps > 0$, there
	exists a family of fingerprints $\mathcal{S}(n,\eps)$ and
	$\mathcal{G}(n,\eps)$ of subgraphs of $K_n$ satisfying the following
	properties:
	\begin{enumerate}[label=(\alph*)]
		\item $|\mathcal{G}| \leq \exp(C(\eps)n^{3/2}\log n).$
		\item For every $G \in \mathcal{G}$, $k_3(G) \leq \eps n^3.$
		\item For every triangle free graph $H$ on $n$ vtx, there exists $G
			\in \mathcal{G}$ with $H \subset G$.
	\end{enumerate}
\end{theorem}

\begin{proof}
	We will iteratively apply our lemma
	[\ref{lemma:hypergraph_containers}] until we are left with our desired family.
	What degree properties can we say about $\mathcal{H} =
	\mathcal{H}_{K_3}(K_n)$? It's simple to note that $e(\mathcal{H}) =
	\binom{n}{3}$,
	$d = \Delta_1(\mathcal{H}) = n-2$ and $\Delta_2(\mathcal{H}) \leq 1$.
 So, the first time, we can apply the lemma with $\tau = 1/\sqrt{n}$.

	Let $\mathcal{C} = \{E(K_n)\}$, while there is an element $E(G)$ in
	$\mathcal{C}$ with more than $\eps n^3$ triangles, we apply the
	container lemma on $G$, and substitute
	$G$ by its containers in $\mathcal{C}$. We know the maximum degree
	condition is the same, and because $k_3(G) \geq \eps n^3$, we know
	that the avg. degree in our
	container lemma is at least $\eps n$. So we can always repeat this
	process with $c = 1$ and $\tau = 1/\eps \sqrt{n}$.

	How many steps can we do? Well after $t$ iterations, each container
	has size at most $(1-\delta)^t\binom{n}{2}$ yielding a maximum of
	$(1-\delta)^t n^3$ triangles.
	So we must have that $(1-\delta)^t \geq \eps$, and therefore $t \leq
	\log(1/\eps)/\delta$. So how many containers do we end up with? Something like
	\[
		\binom{n^2}{n^{3/2}}^t = n^{O(n^{3/2})} = \exp(O(n^{3/2}\log n)).
	\]
	We can even bound the sizes of the fingerprints by $t \tau
	\binom{n}{2} = Cn^{3/2}$.
	% Let $\mathcal{C}_0 = \{\mathcal{H}_{K_3}(K_n)\}$. What
\end{proof}

\subsection{Some cool applications!}

\subsubsection{Counting triangle-free graphs}

We will try to prove the following theorem:
\begin{theorem}
	\label{containers:far_bipartite}
	Almost all triangle free graphs with $n$ vtxs and $m \gg n^{3/2}$
	edges are $o(m)$ close to being bipartite.
\end{theorem}

To prove it we will need an estability result (basically
Erdös-Simonitivs) that states
\begin{theorem}
	If $G$ is $t$-far from being bipartite, then
	\[
		k_3(G) \geq \frac{n}{6}(e(G) + t - \frac{n^2}{4})
	\]
\end{theorem}

\begin{proof}
	(of \ref{containers:far_bipartite}) Let $X(n,m,\alpha)$ be the
	number  of  triangle free graphs with $m$ edges which are $\alpha
	m$-far from being bipartite.
	We will upperbound $X$ counting by counting inside the containers.
	So suppose we applied [\ref{trm:container_for_triangles}] with
	$\eps>0$, which yielded $\mathcal{G}$.
	We have
	\[
		X(n,m,\alpha) \leq \sum_{G \in \mathcal{G}} \#\{ K_3 \not \subset H
		\subset G,\; e(H) = m,\; H\,\alpha m\text{-far from being bip.}\}.
	\]
	Now fix some parameter $\delta > 7\eps$ we choose later, we know
	from the stability result that either $G$ is $\delta n^2$ close to
	being bipartite or
	\[
		e(G) \leq \frac{n^2}{4} + 6\eps n^2 - \delta n^2 \leq \frac{n^2}{4}
		- \eps n^2.
	\]
	So let's separate our counting depending on the container and bound
	the previous sum
	\[
		X(n,m,\alpha) \leq \sum_{\substack{G \in \mathcal{G}\\e(G) \leq
		(1/4 - \eps)n^2}} \binom{e(G)}{m}\quad
		+ \quad\sum_{\substack{G \in \mathcal{G}\\G\, \delta-\text{close to
		bip.}}} \binom{\delta n^2}{\alpha m} \binom{n^2/4}{(1-\alpha)m}
	\]
	where the first inequality follows from taking all $m$ subgraphs of
	$G$ when $G$ has few edges and the second follows from choosing the
	$\alpha m$ edges
	that make $H$ far from being bipartite. A bit of manipulation shows
	\begin{align*}
		X(n,m,\alpha) &\leq \sum_{G \in \mathcal{G}} (1-4\eps)^m
		\binom{n^2/4}{m} + \binom{\delta n^2}{\alpha
		m}\binom{n^2/4}{m}\bigg(\frac{n^2}{4}\bigg)^{-\alpha m}\\
		&\leq n^{Cn^3/2}e^{-cm} \binom{n^2/4}{m} \to o(1)\binom{n^2/4}{m}
	\end{align*}
	when $m \gg n^{3/2}\log n$.

	What if we want to remove the $\log n$? Then we do the same thing
	but we use our fingerprints, we sum over them and remember that each
	$S$ must be cointained in $H$.
	Taking that into consideration the $\log n$ vanishes.

	\begin{align*}
		X(n,m,\alpha) \leq \sum_{|S| = 0}^{Cn^{3/2}}
		\binom{n^2/2}{|S|}\bigg(\binom{(1/4-\eps)n^2}{m - |S|} +
		\binom{n^2/4}{(1-\alpha/2)m - |S|}\bigg)
	\end{align*}
	and after some more careful computation we win.
\end{proof}

\subsubsection{When does Gnp arrows a triangle?}

The following is a very surprising theorem (and beautiful one as well).
\begin{theorem}
	\label{containers:arrows_k3}
	Let fix $r \in \N$ and let $p \gg n^{-1/2}$, then $G(n,p) \to_r K_3$ w.h.p.
\end{theorem}
As always, we need some kind of saturation result. We will actually
show that $\Pr(G(n,p) \not \to K_3) \to 0$. So what does this event really mean?
If $G(n,p) \not \to_r K_3$, then there exists a partition (or
coloring) of the edges of $G(n,p) = H_1 \cup H_2 \dots \cup H_r$,
such that each $H_i$ is $K_3$-free.
Now an easy double counting yields the following lemma:
\begin{lemma}
	Let $G$ be a graph and $\chi$ an r-coloring of its edges. Then if
	\[
		e(G) \geq (1-\delta)\binom{n}{2}
	\]
	then $\chi$ contains $\delta n^3$ monochromatic triangles.
\end{lemma}
And we are ready to apply our containers.
\begin{proof}
	(of \ref{containers:arrows_k3}) We upperbound again on our
	containers, let $C_i$ be the containers for $H_i$. We know $E(K_n) =
	F \cup C_1 \dots \cup C_r$
	where $k_3(C_i) \leq \eps n^3$ for each $i$ and $F$ is the
	complement of the union.
	By the previous lemma, it must be that $e(F) \geq \eps r
	\binom{n}{2}$, and this means that, if $G(n,p) \subset C_1 \cup
	\dots \cup C_r$,
	it is avoiding a positive density set of edges. We find that
	(maximizing by the last term)
	\begin{align*}	
		\Pr(G(n,p) \not \to K_3) &\leq \sum_{S_1 \cup \dots \cup S_r}
		\Pr((S_1 \cup \dots S_r) \subset G(n,p) \subset (C_1 \cup \dots \cup C_r))\\
		&\leq \sum_{t = 0}^{Crn^{3/2}} \binom{n^2/2}{t}p^t(1-p)^{\delta
		n^2} \leq e^{-p\delta n^2}\bigg(\frac{epn^2}{Crn^{3/2}}\bigg)^{Crn^{3/2}}\\
		&= \bigg(\frac{e^{-\delta p\sqrt{n}}p\sqrt{n}}{D}\bigg)^{Crn^{3/2}}
	\end{align*}
	and it is clear that if $p\sqrt{n} \to \infty$, then the probability
	is tends to 0.

\end{proof}
