¥appendix
¥section{(Sub)Exponential Tailbounds}

I just took these from my notes in random matrices.

¥begin{lemma}
	¥label{lemma:hoeff}
	(Hoeffding's Lemma) Let $X$ be a scalar r.v taking values in
	$[a,b]$. Then for any $t > 0$
	¥begin{equation}
		¥label{lemma:hoef:eq1}
		¥E e^{tX} ¥leq e^{t¥Ex X}(1 + O(t^2¥sigma^2¥exp(O(t(b-a))))).
	¥end{equation}
	In particular
	¥begin{equation}
		¥label{lemma:hoef:eq2}
		¥E e^{tX} ¥leq e^{t¥E X}¥exp(O(t^2(b-a)^2)).
	¥end{equation}
¥end{lemma}

¥begin{proof}
	Normalize $X$ as to $¥Ex(X) = 0$ and $b - a = 1$. And consider the
	Taylor expansion of $e^{tX}$,
	¥[
		e^{tX} = 1 + tX + O(t^2X^2¥exp(O(t)))
	¥]
	taking expectation yields the first result. The second one follows
	from assymptotics of the exponential function.
¥end{proof}

¥begin{theorem}
	¥label{trm:McDiamird}
	(McDiamird's Theorem) Let $(R_i)_{i=1}^n$ be a tuple of ranges
	and $X = (X_i)_{i=1}^{n}$ a random vector taking values in
	their product.
	Let $F: ¥prod_{i=1}^n R_i ¥to ¥R$ be a coordinate-wise bounded
	function (this is a strong Lipchitz condition). That is changing
	any coordinate $x_i$ may change at most $c_i$ the total value.
	Then
	¥[
		¥Pr( |F(X) - ¥Ex F(X)| > ¥lambda ¥sigma) ¥leq C ¥exp(-c ¥lambda^2)
	¥]
	for some $C,c > 0$ and $¥sigma^2 = ¥sum_{i=1}^n c_i^2$.
¥end{theorem}

¥begin{proof}
	By simmetry, we may only consider $¥Pr(F(X) - ¥E F(X) > ¥lambda
	¥sigma)$. We turn to Hoeffding's Lemma [¥ref{lemma:hoeff}],
	consider
	¥[
		¥Ex ¥exp(tF(X)) = ¥Ex ¥big[ ¥Ex[¥exp(tF(X))| X_1, ¥dots X_{n-1}] ¥big],
	¥]
	we should try to understand $¥Ex[¥exp(tF(X))| X_1, ¥dots X_{n-1}]$.
	We have that
	¥[
		¥Ex[¥exp(tF(X))| X_1, ¥dots X_{n-1}] = ¥Ex[¥exp(t(F(X) -
		¥Ex(F)))¥exp(t¥Ex F(X))| X_1, ¥dots X_{n-1}]
	¥]
	applying (¥ref{lemma:hoef:eq2}) to the first product, (here we are
	conditioning on $X_1, ¥dots X_n$)
	¥[
		¥Ex[¥exp(t(F(X) - ¥Ex(F)))¥exp(t¥Ex F(X))| - ] ¥leq
		¥exp(O(t^2c_n^2))¥Ex[¥exp(t¥Ex F(X)) | -]
	¥]
	and we can iteratively apply the theorem to the new coordinate-wise
	bounded function
	¥[
		G(X_1, ¥dots X_{n-1}) = ¥Ex[F(X)|X_1, ¥dots X_{n-1}].
	¥]
	Notice $G$ has the same coefficients as $F$, because
	¥begin{align*}
		|G(y, x_2 ¥dots, x_{n-1}) - G(z, x_2 ¥dots, x_{n-1})| &= ¥bigg|
		¥int_{R_n} F(y,¥dots,x_n) dx_n - ¥int_{R_n} F(z,¥dots,x_n) dx_n¥bigg|¥¥
		&= ¥bigg| ¥int_{R_n} F(y,¥dots,x_n) - F(z,¥dots,x_n) dx_n¥bigg| < c_1,
	¥end{align*}
	and here we can take any other coordinate instead of the first.
¥end{proof}

¥begin{corollary}
	¥label{trm:hoeffding_ineq}
	(Hoeffding's inequality) Let $(X_i)_{i=1}^n$ be real r.v's taking
	value in bounded intervals $[a_i,b_i]$. Then, if $S = ¥sum_{i ¥leq
	n} X_i$, we have
	¥[
		¥Pr( |S  - ¥Ex S| > ¥lambda¥sigma) < C¥exp(-c ¥lambda^2)
	¥]
	for some $C,c > 0$, where $¥sigma^2 = ¥sum_{i ¥leq n} (b_i-a_i)^2$.
¥end{corollary}
¥begin{proof}
	Apply [¥ref{trm:McDiamird}] with $F(X_1, ¥dots X_n) = X_1 + ¥dots + X_n$.
¥end{proof}

¥begin{theorem}
	¥label{trm:chernoff} (Chernoff) Let $(X_i)_{i=1}^n$ be independent
	scalar random variables with $|X_i| < K$ a.s,
	with mean $¥mu_i$, and variance $¥sigma_i^2$. Then for any $¥lambda
	> 0$, one has
	¥[
		¥Pr(|S_n - ¥mu| ¥geq ¥lambda ¥sigma) ¥leq C¥exp(-c ¥min(¥lambda^2,
		¥frac{¥lambda ¥sigma}{K}))
	¥]
	for some absolute $C,c > 0$, where $S_n = ¥sum_{i=1}^{n} X_i$, $¥mu
	= ¥sum_{i=1}^{n} ¥mu_i$ and $¥sigma^2 = ¥sum_{i=1}^{n} ¥sigma_i^2$.
¥end{theorem}

¥begin{proof}
	The proof is again the use of the exponential moment with lemma
	[¥ref{lemma:hoeff}]. Suppose we normalize  each variable
	so that $K ¥leq 1$ and $¥mu_i = 0$, then $¥mu = 0$ and for any $0
	¥leq t ¥leq 1$,
	¥[
		¥Ex(¥exp(tS_n)) = ¥prod_{i=1}^{n} ¥Ex(¥exp(tX_i)) = ¥prod_{i=1}^n
		(1 + O(t^2¥sigma_i^2¥exp(O(t)))).
	¥]
	using the usual $1 + x ¥leq e^x$ ineq, we find
	¥[
		¥Ex(¥exp(tS_n)) ¥leq ¥exp(t¥Ex(S_n))¥exp(O(t^2)¥exp(O(t))¥sigma^2)
		= ¥exp(O(t^2¥sigma^2)),
	¥]
	now we use Markov and optimize on $t$ following the restraint on $0
	¥leq t ¥leq 1$. (Suppose the hidden constant is $D$)
	we have that
	¥[
		¥Pr(S_n ¥geq ¥lambda¥sigma) ¥leq ¥exp(D t^2¥sigma^2 - t¥lambda¥sigma)
	¥]
	which yields the result after optimizing $t$ following the
	constraints. The other inequality is equivalent.

¥end{proof}

¥begin{corollary}
	¥label{lemma:expec_binomial_bounds}
	(Tailbounds for the Binomial distribution) Let $B(n,p)$ be the
	binomial distribution.
	We have that
	¥[
		¥Pr(|B(n,p) - np| ¥geq ¥delta np) ¥leq C¥exp(-c ¥delta^2 (np)^2)
	¥]
	for some absolute $C,c$.
¥end{corollary}
¥begin{proof}
	Note that
	¥[
		¥Pr(|B(n,p) - np| ¥geq ¥lambda ¥sqrt{np}) ¥leq ¥Pr(|B(n,p) - np|
		¥geq ¥lambda ¥sqrt{np(1-p)}) ¥leq C¥exp(-c¥lambda^2)
	¥]
	by ¥ref{trm:chernoff}. Applying it with $¥lambda = ¥delta
	¥sqrt{np}$ yields the result.
¥end{proof}

¥begin{theorem}
	(Azuma's Inequality) Let $X_1, X_2, ¥dots X_n$ be a sequence scalar
	r.v's with $|X_i| ¥leq 1$ almost surely. Assume also
	that we have the martingale difference property
	¥[
		¥E(X_i | X_1, ¥dots, X_{i-1}) = 0
	¥]
	almost surely for all $i = 1,¥dots,n$. Then for any $¥lambda > 0$,
	the sum $S_n := X_1 + ¥dots + X_n$ satisfies
	¥[
		¥Pr(|S_n| ¥geq ¥lambda ¥sqrt{n}) ¥leq C¥exp(-c¥lambda^2)
	¥]
	for some absolute $C,c > 0$.
¥end{theorem}

¥begin{proof}
	We prove this by calculating exponential moments and using
	Hoeffding's Lemma again. Note that $S_n = S_{n-1} + X_n$ so that
	¥[
		¥Ex(e^{tS_n}) = ¥Ex(e^{tS_{n-1}}e^{tX_n}).
	¥]
	By taking the expectation of conditioning on the first $n-1$
	values, we get that
	¥[
		¥Ex(e^{tS_n}) = ¥Ex ¥Ex (e^{tS_{n-1}}e^{X_n} | X_1, ¥dots,
		X_{n-1}) =¥Ex [e^{tS_{n-1}}¥Ex(e^{tX_n} | X_1, ¥dots, X_{n-1})]
	¥]
	Now, by ¥ref{lemma:hoeff}, we have that $¥Ex(e^{tX_n} | X_1, ¥dots,
	X_{n-1}) ¥leq e^{O(t^2)}$, so by iterating we get the result.

¥end{proof}
